---
title: "Links for Teaching with R"
author: "Daniel Kaplan"
date: "April 7, 2014"
output: html_document
---

```{r include=FALSE}
library( mosaic )
```

## Monday March 7, 2014

I usually have an online syllabus with a link for each day's notes.

In that page, I can put whatever resources are needed for that class day.

```{r}
library( mosaic )
```

## Today's Resources

The [**scaffold** for the in-class activity](https://raw.githubusercontent.com/dtkaplan/NIMBIOS/master/inst/NIMBios-Slides/Rsquared-scaffold.Rmd).  Cut and paste this into an RStudio editor window.  Then customize it and hand it in on RPubs.

## Shiny

* [Langrange Multipliers](http://glimmer.rstudio.com/mosaic/LagrangeMultiplier/)
* [Predicting the 2012 Presidential Election](http://glimmer.rstudio.com/mosaic/Election2012/)
* [Geometry of Confidence Intervals](http://glimmer.rstudio.com/mosaic/CIgeometry/)
* [Bayes App](http://glimmer.rstudio.com/mosaic/BayesApp/) written by a student, Ross Peterson, who took both the Bayesian Statistics and the Scientific Programming courses in Fall 2013. 

-------------------------

## The Basement 

Things you don't really need.

* The Google Spreadsheet: [editing link](https://docs.google.com/spreadsheet/ccc?key=0Am13enSalO74dEtyWGxpWWFsN3Z0OUlZNG5xYmRVWWc&usp=sharing) and [fetching](https://docs.google.com/spreadsheet/pub?key=0Am13enSalO74dEtyWGxpWWFsN3Z0OUlZNG5xYmRVWWc&single=true&gid=0&output=csv)

To fetch, cut and paste the following from this page into R

```{r}
ourData <- fetchGoogle("https://docs.google.com/spreadsheet/pub?key=0Am13enSalO74dEtyWGxpWWFsN3Z0OUlZNG5xYmRVWWc&single=true&gid=0&output=csv")
```
